---
title: "homework 6"
author: "lauren franks, lef2147"
date: "11/20/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readr)
library(modelr)
```

Load and Tidy Data
* Load and clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).
```{r}
bw_data = read_csv("birthweight.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    babysex = as.factor(babysex),
    frace = as.factor(frace),
    malform = as.factor(malform),
    mrace = as.factor(mrace)
  )
  
skimr::skim(bw_data)
```

Visualizing Birthweight
  - slight left skew
```{r}
bw_data %>% 
  ggplot(aes(bwt)) + geom_histogram()
```




Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. Describe your modeling process and show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.
```{r}
fit = lm(bwt ~ ., data = bw_data)

#car::vif(fit)
# error given here indicates that there is extremely high collinearity going on, so we will fit a model with less variables and reassess multicollinearity 

# there are three variables that describe number of previous births (parity, pnumlbw, pnumsga) and four variables that explain mom's weight (ppbmi, ppwt, wtgain, delwt) we will choose only one/two variable in each category to keep and reassess multicollinearity
fit_1 = lm(bwt ~ babysex + frace + malform + bhead + blength + fincome + gaweeks + menarche + mheight + momage + parity + smoken + ppwt + wtgain, data = bw_data)

car::vif(fit_1)
# there is no multicollinearity present, we can now evaluate the model 

summary.lm(fit_1)

summary(fit_1)$adj.r.squared

# remove malform1 because the standard error is extremely high
fit_2 = lm(bwt ~ babysex + frace + bhead + blength + fincome + gaweeks + menarche + mheight + momage + parity + smoken + ppwt + wtgain, data = bw_data)

summary.lm(fit_2)

# remove momage
fit_3 = lm(bwt ~ babysex + frace + bhead + blength + fincome + gaweeks + menarche + mheight + parity + smoken + ppwt + wtgain, data = bw_data)

summary.lm(fit_3)

# remove menarche
fit_4 = lm(bwt ~ babysex + frace + bhead + blength + fincome + gaweeks + mheight + parity + smoken + ppwt + wtgain, data = bw_data)

summary.lm(fit_4)

# remove fincome
fit_5 = lm(bwt ~ babysex + frace + bhead + blength + fincome + gaweeks + mheight + parity + smoken + ppwt + wtgain, data = bw_data)

summary.lm(fit_5)

summary(fit_1)$adj.r.squared
summary(fit_2)$adj.r.squared
summary(fit_3)$adj.r.squared
summary(fit_4)$adj.r.squared
summary(fit_5)$adj.r.squared
```

Using Model 5
```{r}

modelr::add_residuals(bw_data, fit_5) %>% 
  ggplot(aes(bwt, resid)) + geom_violin()

modelr::add_residuals(bw_data, fit_5) %>% 
  ggplot(aes(bwt, resid)) + geom_point()

fit_6 = lm(bwt ~ babysex + smoken + ppwt + wtgain + fincome + gaweeks + frace + mheight + blength + parity, data = bw_data)

modelr::add_residuals(bw_data, fit_6) %>% 
  ggplot(aes(bwt, resid)) + geom_point()

# fit 6 has a less pronounced curved pattern

# modeling both predicted values against redsiduals
modelr::add_predictions(bw_data, fit_6) %>% 
  mutate(
    resid = pred - bwt
  ) %>% 
  ggplot(aes(pred, resid)) + geom_point()

# aside from a few outliers, the residuals are evenly scattered about 0
```


Comparing model to two other models:
* One using length at birth and gestational age as predictors (main effects only)
* One using head circumference, length, sex, and all interactions (including the three-way interaction) between these
```{r}
model_1 = lm(bwt ~ blength + menarche, data = bw_data)

summary.lm(model_1)

model_2 = lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + babysex*blength, data = bw_data)

summary.lm(model_2)
```


Make this comparison in terms of the cross-validated prediction error; use crossv_mc and functions in purrr as appropriate.
```{r}
train = sample_frac(bw_data, size = 0.8)
test = anti_join(bw_data, train)

cv_df = crossv_mc(bw_data, 100)

cv_df =
  cv_df %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )

cv_results = cv_df %>% 
  mutate(
    model_1 = map(.x = train, ~lm(bwt ~ blength + menarche, data = .x)),
    model_2 = map(.x = train, ~lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + babysex*blength, data = .x)),
    fit_6 = map(train, ~lm(bwt ~ babysex + smoken + ppwt + wtgain + fincome + gaweeks + frace + mheight + blength + parity, data = bw_data)),
    rmse_1 = map2(.x = model_1, .y = test, ~rmse(.x, .y)),
    rmse_2 = map2(.x = model_2, .y = test, ~rmse(.x, .y)),
    rmse_3 = map2(fit_6, test, ~rmse(model = .x, data = .y))
  )


# cv_results %>% select(starts_with("rmse")) %>% 
#   pivot_longer(
#     everything(),
#     names_to = "model", 
#     values_to = "rmse",
#     names_prefix = "rmse_") %>% 
#   mutate(model = fct_inorder(model)) %>% 
#   ggplot(aes(x = model, y = rmse)) + geom_violin()
```


Note that although we expect your model to be reasonable, model building itself is not a main idea of the course and we don’t necessarily expect your model to be “optimal”.












